---
title: "pilot"
output:
  pdf_document: default
  html_document:
df_print: paged
header-includes:
 \usepackage{booktabs}
 \usepackage{longtable}
 \usepackage{array}
 \usepackage{multirow}
 \usepackage{wrapfig}
 \usepackage{float}
 \floatplacement{figure}{H}

---

# Pilot

```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H', 
                      echo=TRUE, warning=FALSE, message=FALSE)
```

# Includes

```{r includes}
library(tidyverse)
library(estimatr)
library(ggplot2)
library(readr)
library(broom)
library(viridis)


th <- theme_classic()
theme_set(th)
```

# Load behavioral data and covariates

```{r load}
csv_dir = "pixeldiffs-equated"

keeps <- c('RT', 'TrialName', 'Response', 'WID', 'presentation_time',
           'type', 'same_gt', 'same_response', 'idx', 'flipped')
data <-
  list.files(path=file.path("behavioral_data", csv_dir), pattern = "*.csv", 
                            full.names = TRUE, recursive = TRUE) %>% 
  map_df(~read_csv(., show_col_types = FALSE)) %>%
  subset(select = keeps) %>%
  rename(name = TrialName, 
         time.presentation = presentation_time,
         scene = idx) %>%
  mutate(correct = !xor(same_gt, same_response))

# keep 150, 250, or 350 presentation times
data <- data %>%
  filter(time.presentation %in% c(150,250,350))

drops <- c('stim_time', 'mask_time')
covariates <- read_csv(file.path("covariates", csv_dir, "movies_output.csv"))
covariates <- covariates[,!(names(covariates) %in% drops)]

unfiltered <- data %>%
  left_join(covariates, by = c("name"))

if (csv_dir == "pixeldiffs-unequated") {
  # exclude 09_...mp4 as it has super large pixel distance 
  unfiltered <- unfiltered %>%
    filter(!grepl("09_", name))
}
```
## Mark trials with suspicious z-scores
```{r}
# unfiltered <- unfiltered %>%
#   mutate(RT.zscore = scale(RT)) %>%
#   mutate(high.zscore = abs(RT.zscore >= 3))
RT.zscore <- ave(unfiltered$RT, unfiltered$WID, FUN = scale)
unfiltered$high.zscore = abs(RT.zscore) >= 3 | unfiltered$RT > 6000
```

# Visualizations of the raw data

## Reaction times for each trial per subject
```{r}
unfiltered %>%
  ggplot(aes(x = RT, y = WID)) +
  geom_point(aes(col = high.zscore))

unfiltered %>%
  filter(high.zscore) %>%
  group_by(WID) %>%
  summarize(n.excluded = n()) %>%
  print()
```

# Filter data
Exclude subjects whose number of hits in the different trials are less than 1.5 times the number of false positives in the same trials. Then, exclude trials whose reaction times wrt to an individual subject are too extreme, i.e., abs(zscore) >= 3.

```{r}
by_subj <- unfiltered %>%
  group_by(WID)

hr_by_subj <- by_subj %>%
  filter(!same_gt) %>%
  summarise(hr = mean(correct))

fp_by_subj <- by_subj %>%
  filter(same_gt) %>%
  summarise(fp = 1.0 - mean(correct))

subject_performance <- hr_by_subj %>%
  left_join(fp_by_subj, by = "WID")

passed_subjects <- subject_performance %>%
  filter(hr > 1.5*fp)

filtered <- passed_subjects %>%
  left_join(unfiltered, by = "WID") %>%
  filter(!high.zscore)

filtered %>%
  group_by(time.presentation, flipped) %>%
  summarize(n.good_subjects = n_distinct(WID)) %>%
  print()
```

# Analysis: hit rate by trial

## Dataframe preparation
```{r}
by_trial <- filtered %>%
  group_by(scene, type, time.presentation, flipped)

# hit rate
hr_by_trial <- by_trial %>%
  filter(!same_gt) %>%
  summarize(rate.hit = mean(correct),
            rate.hit.se = sd(correct) / sqrt(length(correct)),
            dist.pixel = mean(`pixel-distance`),
            dist.light = mean(`light-distance`),
            dist.mesh = mean(`mesh-distance`),
            dist.pose = mean(`posevec-distance`),
            dist.shape = mean(`shapevec-distance`))

# false alarm rate
fa_by_trial <- by_trial %>%
  filter(same_gt) %>%
  summarize(rate.fa = 1 - mean(correct)) %>%
  subset(select = -c(type))

# dprime
dprime_by_trial <- hr_by_trial %>%
  left_join(fa_by_trial, by = c("scene", "time.presentation", "flipped")) %>%
  mutate(d.prime = qnorm(rate.hit*0.999 + 0.0001) - qnorm(rate.fa*0.999 + 0.0001))
```

## hit rate

### plots
```{r}
fa_by_trial %>%
  group_by(time.presentation, flipped) %>%
  summarize(mean = mean(rate.fa)) %>%
  ggplot(aes(x = time.presentation, y = mean)) +
  geom_bar(stat = "identity") +
  facet_wrap( ~ flipped) +
  ggtitle("False alarm rate as a function of presentation time")

bar_plot <- function(.data, y, ylim, ylabel){
  .data %>%
    group_by(type, flipped) %>%
    summarize(mean = mean(!! sym(y)),
              se = sd(!! sym(y)) / sqrt(length(!! sym(y))),
              ci = abs(qnorm((1-0.95)/2))* sd(!! sym(y))/sqrt(length(!! sym(y)))) %>%
    ggplot(aes(x = type)) +
    geom_bar(aes(y=mean, fill = type), stat="identity") +
    ylim(ylim) +
    geom_linerange(aes(ymin=mean-ci, ymax=mean+ci)) +
    labs(x="Cause of visual change", y=ylabel) + 
    scale_fill_brewer(palette = "Accent") +
    theme(legend.position = "none") + 
    scale_x_discrete(breaks=c("different-light","different-posture"),
        labels=c("light", "posture"))
}

w = 40 
h = 66.73
svg(file='upright_accuracy.svg',width=w*0.039370,height=h*0.039370,pointsize=7)
hr_by_trial %>%
  filter(!flipped) %>%
  bar_plot("rate.hit", c(0, 1), "Detection Accuracy")
dev.off()

svg(file='upright_dprime.svg',width=w*0.039370,height=h*0.039370,pointsize=7)
dprime_by_trial %>%
  filter(!flipped) %>%
  bar_plot("d.prime", c(0,2), "Detection Sensitivity d`") 
dev.off()

svg(file='flipped_accuracy.svg',width=w*0.039370,height=h*0.039370,pointsize=7)
hr_by_trial %>%
  filter(flipped) %>%
  bar_plot("rate.hit", c(0, 1), "Detection Accuracy")
dev.off()

svg(file='flipped_dprime.svg',width=w*0.039370,height=h*0.039370,pointsize=7)
dprime_by_trial %>%
  filter(flipped) %>%
  bar_plot("d.prime", c(0,2), "Detection Sensitivity d`") 
dev.off()

hr_by_trial %>%
  ggplot(aes(x = dist.pixel, y = rate.hit, col = type)) +
  geom_text(aes(label = scene), 
            size = 2) #+
  #facet_wrap( ~ time.presentation)
```

```{r}
corr_plot <- function(.data, x, y, xlabel, ylabel) {
  .data %>%
    ggplot(aes(x = !! sym(x), y = !! sym(y), fill = type)) + 
    geom_point(aes(color = type)) + 
    geom_smooth(method = "lm") +
    labs(x=xlabel, y=ylabel) + 
    scale_fill_brewer(palette = "Accent") +
    scale_color_brewer(palette = "Accent") +
    theme(legend.position = "none") + 
    # theme(axis.title.x = element_blank(),
    #       axis.title.y = element_blank()) +
    ylim(c(0,3))
}

hr_by_trial %>% 
  filter(!flipped) %>%
  corr_plot("dist.pixel", "rate.hit", "pixel distance", "accuracy")
hr_by_trial %>%
  filter(!flipped) %>%
  corr_plot("dist.light", "rate.hit", "light distance", "accuracy")
hr_by_trial %>%
  filter(!flipped) %>%
  corr_plot("dist.mesh", "rate.hit", "mesh distance", "accuracy")

w = 50
h = 38
svg(file='upright_pixel.svg',width=w*0.039370,height=h*0.039370,pointsize=7)
dprime_by_trial %>% 
  filter(!flipped) %>%
  corr_plot("dist.pixel", "d.prime", "pixel distance", "d`")
# ggsave("upright_pixel.pdf", width = 43, height = 31, units = "mm")
dev.off()
svg(file='upright_light.svg',width=w*0.039370,height=h*0.039370,pointsize=7)
dprime_by_trial %>%
  filter(!flipped) %>%
  corr_plot("dist.light", "d.prime", "light distance (deg)", "d`")
dev.off()
svg(file='upright_mesh.svg',width=w*0.039370,height=h*0.039370,pointsize=7)
dprime_by_trial %>%
  filter(!flipped) %>%
  corr_plot("dist.mesh", "d.prime", "mesh distance", "d`")
dev.off()

svg(file='flipped_pixel.svg',width=w*0.039370,height=h*0.039370,pointsize=7)
dprime_by_trial %>% 
  filter(flipped) %>%
  corr_plot("dist.pixel", "d.prime", "pixel distance", "d`")
# ggsave("upright_pixel.pdf", width = 43, height = 31, units = "mm")
dev.off()
svg(file='flipped_light.svg',width=w*0.039370,height=h*0.039370,pointsize=7)
dprime_by_trial %>%
  filter(flipped) %>%
  corr_plot("dist.light", "d.prime", "light distance (deg)", "d`")
dev.off()
svg(file='flipped_mesh.svg',width=w*0.039370,height=h*0.039370,pointsize=7)
dprime_by_trial %>%
  filter(flipped) %>%
  corr_plot("dist.mesh", "d.prime", "mesh distance", "d`") +
  ylim(c(0,3))
dev.off()

# dprime_by_trial %>%
#   group_by(time.presentation, type) %>%
#   summarize(mean = mean(dist.pixel),
#             sd = sd(dist.pixel)) 

```

## model-free statistics

```{r}
t.test(filter(hr_by_trial, !flipped)$rate.hit, mu = 0.5)
t.test(filter(hr_by_trial, !flipped & type == "different-posture")$rate.hit, mu = 0.5)
t.test(filter(hr_by_trial, !flipped & type == "different-light")$rate.hit, mu = 0.5)
t.test(filter(hr_by_trial, !flipped & type == "different-posture")$rate.hit, 
       filter(hr_by_trial, !flipped & type == "different-light")$rate.hit,
       paired = TRUE, alternative = "two.sided")
t.test(filter(dprime_by_trial, !flipped & type == "different-posture")$d.prime, 
       filter(dprime_by_trial, !flipped & type == "different-light")$d.prime,
       paired = TRUE, alternative = "two.sided")
```

```{r}
t.test(filter(hr_by_trial, flipped)$rate.hit, mu = 0.5)
t.test(filter(hr_by_trial, flipped & type == "different-posture")$rate.hit, mu = 0.5)
t.test(filter(hr_by_trial, flipped & type == "different-light")$rate.hit, mu = 0.5)
t.test(filter(hr_by_trial, flipped & type == "different-posture")$rate.hit, 
       filter(hr_by_trial, flipped & type == "different-light")$rate.hit,
       paired = TRUE, alternative = "two.sided")
t.test(filter(dprime_by_trial, flipped & type == "different-posture")$d.prime, 
       filter(dprime_by_trial, flipped & type == "different-light")$d.prime,
       paired = TRUE, alternative = "two.sided")
```

```{r}
t.test(filter(dprime_by_trial, flipped & type == "different-posture")$d.prime, 
       filter(dprime_by_trial, !flipped & type == "different-posture")$d.prime,
       paired = TRUE, alternative = "two.sided")

t.test(filter(dprime_by_trial, flipped & type == "different-light")$d.prime, 
       filter(dprime_by_trial, !flipped & type == "different-light")$d.prime,
       paired = TRUE, alternative = "two.sided")
```


```{r}
hr_by_trial %>%
  nest(data = -time.presentation) %>% 
    mutate(model = map(data, ~lm(rate.hit~dist.pixel + type, data = .)), 
           tidied = map(model, tidy)) %>% 
    unnest(tidied)

hr_by_trial %>%
  with(lm(rate.hit ~ dist.pixel + dist.light + dist.mesh + type)) %>%
  summary()

dprime_by_trial %>%
  with(lm(d.prime ~ dist.pixel + dist.light + dist.mesh + type)) %>%
  summary()
```
```{r}
dprime_by_trial %>%
  filter(type == "different-posture") %>%
  with(lm(d.prime ~ dist.mesh)) %>%
  summary()

dprime_by_trial %>%
  filter(type == "different-posture") %>%
  with(lm(rate.hit ~ dist.mesh)) %>%
  summary()
```

```{r}
dprime_by_trial %>%
  filter(type == "different-posture" & !flipped) %>%
  with(lm(d.prime ~ dist.pixel)) %>%
  summary()
dprime_by_trial %>%
  filter(type == "different-posture" & !flipped) %>%
  with(lm(d.prime ~ dist.mesh)) %>%
  summary()

dprime_by_trial %>%
  filter(type == "different-light" & !flipped) %>%
  with(lm(d.prime ~ dist.pixel)) %>%
  summary()
dprime_by_trial %>%
  filter(type == "different-light" & !flipped) %>%
  with(lm(d.prime ~ dist.light)) %>%
  summary()
```

```{r}
dprime_by_trial %>%
  filter(type == "different-posture" & flipped) %>%
  with(lm(d.prime ~ dist.pixel)) %>%
  summary()
dprime_by_trial %>%
  filter(type == "different-posture" & flipped) %>%
  with(lm(d.prime ~ dist.mesh)) %>%
  summary()

dprime_by_trial %>%
  filter(type == "different-light" & flipped) %>%
  with(lm(d.prime ~ dist.pixel)) %>%
  summary()
dprime_by_trial %>%
  filter(type == "different-light" & flipped) %>%
  with(lm(d.prime ~ dist.light)) %>%
  summary()
```

```{r}
# hr_by_trial %>%
#   filter(time.presentation == 150) %>%
#   with(lm(rate.hit ~ dist.pixel + type)) %>%
#   summary()
# 
# hr_by_trial %>%
#   filter(time.presentation == 250) %>%
#   with(lm(rate.hit ~ dist.pixel + type)) %>%
#   summary()
# 
# hr_by_trial %>%
#   filter(time.presentation == 350) %>%
#   with(lm(rate.hit ~ dist.pixel + type)) %>%
#   summary()
```

# Analysis with dprime

## Plots
```{r}
dprime_by_trial %>%
  group_by(type, time.presentation) %>%
  summarize(mean = mean(d.prime),
            se = sd(d.prime) / sqrt(length(d.prime))) %>%
  ggplot(aes(x = type)) +
  geom_bar(aes(y=mean), stat="identity") +
  ylim(0, 1) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se)) +
  facet_wrap( ~ time.presentation) + 
  ggtitle("Hit rate as a function of change type")

dprime_by_trial %>% 
  corr_plot("dist.pixel", "d.prime", "pixel distance")
dprime_by_trial %>%
  corr_plot("dist.light", "d.prime", "light distance")
dprime_by_trial %>%
  corr_plot("dist.mesh", "d.prime", "mesh distance")
dprime_by_trial %>%
  corr_plot("dist.pose", "d.prime", "pose-vec distance")
dprime_by_trial %>%
  corr_plot("dist.shape", "d.prime", "shape-vec distance")
```